---
title: "DSS Fall 2017 Seminar 3"
output: html_notebook
authors: Chris & Christy
---


## Dependencies
### New R Packages

There is are three new `R` packages that we'll need for this week. The first, `mapproj`, is the dependency that we were missing at the end of the last session. The second, `RColorBrewer`, and the third, `viridis`, each give us access to a bunch color schemes that are ideal for mapping spatial data. The fourth package, `tidycensus`, makes it easy to download spatial data about St. Louis (and any other city in the U.S.!).

```{r}
install.packages("mapproj")
install.packages("RColorBrewer")
install.packages("viridis")
install.packages("tidycensus")
```

Once they are installed, we can load them with the `library()` function:

```{r}
library(mapproj)       # tools for working with map projections
library(RColorBrewer)  # color ramps
library(viridis)       # color ramps
library(tidycensus)    # accessing spatial data
```

### Other R Packages

You will also need a number of other packages that we've worked with in the past sessions:

```{r}
library(broom)     # tidying data
library(ggplot2)   # plotting data
library(maptools)  # tools for working with spatial data
library(rgdal)     # work with shapefiles
```


## Review - Reading Geometric Data in R
We use the `rgdal` package to read shapefiles into `R`. These data must be in the **top-level** of your `R` project directory. To import our `missouriTract` shapefile, we use the `readOGR()` function. We need to specify `dsn = "."` for each use of the function, and the layer should match the name of your shapefile:

```{r}
import <- readOGR(dsn = ".", layer = "missouriTract")
```

Once imported, these data will appear under the "Values" section of the global environment as an object named `import`. However, they are not in a data frame structure. We need to use the `broom` package's `tidy()` function to disaggregate these data and store them in a data frame:

```{r}
missouriTract <- tidy(import, region = "GEOID")
```



Now, you try importing the data from the `stlTract` shapefile using the `readOGR()` function into the `import` object. Then tidy them using the `tidy()` function and save the output into an object named `stlTract`. You are combining the two code chunk above (skip the `View()` code chunk) into a single code chunk:

```{r}

```

## Plotting Census Tracts

We can create simple maps of just the census tracts using `ggplot2` and the `geom_polygon()` function we introduced above:

```{r}
ggplot() +
  geom_polygon(data = missouriTract, mapping = aes(x = long, y = lat, group = id)) +
  coord_map()
```

Now try plotting just the census tracts in St. Louis using the `stlTract` data:

```{r}

```

